Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-11-28 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
worked more on subsampling

```{r}
hamby173and252_train_sub <- hamby173and252_train %>%
  filter(samesource == FALSE) %>%
  slice(sample(1:length(hamby173and252_train$barrel1), 4500, replace = FALSE)) %>%
  bind_rows(hamby173and252_train %>% filter(samesource == TRUE))
```

@@

```{r}
# Specify the options to use with lime 
hamby224_lime_cases <- data.frame(bin_continuous = c(rep(TRUE, 10), rep(FALSE, 2)),
                                    quantile_bins = c(rep(TRUE, 5), rep(FALSE, 5), rep(TRUE, 2)),
                                    nbins = c(rep(2:6, 2), rep(4, 2)),
                                    use_density = c(rep(TRUE, 10), TRUE, FALSE)) %>%
  mutate(case = 1:length(nbins)) %>%
  select(case, bin_continuous, nbins, quantile_bins, use_density)

# Save the cases
saveRDS(hamby224_lime_cases, "../data/hamby224_lime_cases.rds")
```

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../data/hamby224_lime.rds")) {

  # Run the lime functions for the specified number of bins
  hamby224_all_results <- pmap(.l = as.list(hamby224_lime_cases[,-1]),
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3)

  # Separate the lime and explain function results
  hamby224_lime <- map(hamby224_all_results, function(list) list$lime)
  hamby224_explain <- map_df(hamby224_all_results, function(list) list$explain)

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:dim(hamby224_lime_cases)[1], function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s",
              hamby224_lime_cases$bin_continuous[case],
              hamby224_lime_cases$quantile_bins[case],
              hamby224_lime_cases$nbins[case],
              hamby224_lime_cases$use_density[case]))

  # Save the lime and explain results objects and the case info
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

@@

  # Load in the lime results object
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../data/hamby224_sub_lime.rds")) {

  # Run the lime functions for the specified number of bins
  hamby224_sub_all_results <- pmap(.l = as.list(hamby224_lime_cases[,-1]),
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train_sub %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3)

  # Separate the lime and explain function results
  hamby224_sub_lime <- map(hamby224_sub_all_results, function(list) list$lime)
  hamby224_sub_explain <- map_df(hamby224_sub_all_results, function(list) list$explain)

  # Name the items in the lime list
  names(hamby224_sub_lime) <- map_chr(1:dim(hamby224_lime_cases)[1], function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s",
              hamby224_lime_cases$bin_continuous[case],
              hamby224_lime_cases$quantile_bins[case],
              hamby224_lime_cases$nbins[case],
              hamby224_lime_cases$use_density[case]))

  # Save the lime and explain results objects and the case info
  saveRDS(hamby224_sub_lime, "../data/hamby224_sub_lime.rds")
  saveRDS(hamby224_sub_explain, "../data/hamby224_sub_explain.rds")

  # Load in the lime results object
  hamby224_sub_lime <- readRDS("../data/hamby224_sub_lime.rds")
  hamby224_sub_explain <- readRDS("../data/hamby224_sub_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../data/hamby224_sub_bins.csv")) {

  # Create a list a dataframes with the bin boundaries and bins 
  # for the different evaluations of the lime functions
  hamby224_sub_bin_list <- map(hamby224_sub_lime, create_bin_data)

  # Save the bin boundaries and the bins as separate dataframes
  hamby224_sub_bin_boundaries <- map(hamby224_sub_bin_list, function(m) m$boundaries)
  hamby224_sub_bins <- map(hamby224_sub_bin_list, function(m) m$bins)

  # Save the bin boundaries and bins
  saveRDS(hamby224_sub_bin_boundaries, "../data/hamby224_sub_bin_boundaries.rds")
  saveRDS(hamby224_sub_bins, "../data/hamby224_sub_bins.rds")

  # Load in the bin boundaries and bins
  hamby224_sub_bin_boundaries <- readRDS("../data/hamby224_sub_bin_boundaries.rds")
  hamby224_sub_bins <- readRDS("../data/hamby224_sub_bins.rds")

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = pmap_chr(list(feature = feature, 
                                 feature_value = feature_value,
                                 b_c = bin_continuous,
                                 q_b = quantile_bins,
                                 n_b = nbins,
                                 u_d = use_density),
                            .f = bin_labeller, # bin_labeller is one of my helper functions
                            bin_data = hamby224_bin_boundaries,
                            case_info = hamby224_lime_cases)) %>%
    mutate(feature = factor(feature),
           nbins = factor(nbins),
           feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

@@

  # Reorder the variables of feature_desc and feature_bin for plotting purposes
  hamby224_test_explain <- hamby224_test_explain %>%
    mutate(feature_desc = reorder(feature_desc, strictly_less),
           feature_desc = reorder(feature_desc, feature_number),
           feature_desc = reorder(feature_desc, as.numeric(feature))) %>%
    select(bin_continuous:use_density, case:feature_desc, feature_bin:strictly_less, 
           data, prediction)

@@

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../data/hamby224_sub_test_explain.rds")) {

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_sub_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_sub_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = pmap_chr(list(feature = feature, 
                                 feature_value = feature_value,
                                 b_c = bin_continuous,
                                 q_b = quantile_bins,
                                 n_b = nbins,
                                 u_d = use_density),
                            .f = bin_labeller, # bin_labeller is one of my helper functions
                            bin_data = hamby224_sub_bin_boundaries,
                            case_info = hamby224_lime_cases)) %>%
    mutate(feature = factor(feature),
           nbins = factor(nbins),
           feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

  # Finish creating the strictly less than variable
  hamby224_sub_test_explain$strictly_less[grep("< ", hamby224_sub_test_explain$feature_desc)] <- TRUE

  # Reorder the variables of feature_desc and feature_bin for plotting purposes
  hamby224_sub_test_explain <- hamby224_sub_test_explain %>%
    mutate(feature_desc = reorder(feature_desc, strictly_less),
           feature_desc = reorder(feature_desc, feature_number),
           feature_desc = reorder(feature_desc, as.numeric(feature))) %>%
    select(bin_continuous:use_density, case:feature_desc, feature_bin:strictly_less, 
           data, prediction)

  # Save the combined test and explain data
  saveRDS(hamby224_sub_test_explain, "../data/hamby224_sub_test_explain.rds")

  # Load in the data
  hamby224_sub_test_explain <- readRDS("../data/hamby224_sub_test_explain.rds")

@@

  # Create a data frame with the interesting information relating to the differe
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_lime_comparisons <- hamby224_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density) %>%
    slice(1) %>%
    ungroup() %>%
    select(bin_continuous, quantile_bins, nbins, use_density, 
           set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../data/hamby224_sub_lime_comparisons.rds")) {

  # Create a data frame with the interesting information relating to the differe
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_sub_lime_comparisons <- hamby224_sub_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density) %>%
    slice(1) %>%
    ungroup() %>%
    select(bin_continuous, quantile_bins, nbins, use_density, 
           set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

  # Save the lime comparison data frame
  saveRDS(hamby224_sub_lime_comparisons, "../data/hamby224_sub_lime_comparisons.rds")

  # Load in the lime comparison data frame
  hamby224_sub_lime_comparisons <- readRDS("../data/hamby224_sub_lime_comparisons.rds")

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set)

hamby224_sub_lime_results <- hamby224_sub_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set)
```

```{r}
hamby224_lime_results %>%
  group_by(set) %>%
  arrange(mse) %>%
  slice(1:2)

hamby224_sub_lime_results %>%
  group_by(set) %>%
  arrange(mse) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = mse, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid( ~ set) + 
  theme_bw()

hamby224_sub_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = mse, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid( ~ set) + 
  theme_bw()
```

```{r}
hamby224_lime_results %>%
  group_by(set) %>%
  arrange(desc(ave_r2)) %>%
  slice(1:2)

hamby224_sub_lime_results %>%
  group_by(set) %>%
  arrange(desc(ave_r2)) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = ave_r2, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid( ~ set) + 
  theme_bw()

hamby224_sub_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = ave_r2, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid( ~ set) + 
  theme_bw()
```

@@

```{r}
hamby224_test_explain %>%
  ggplot(aes(x = feature_desc)) + 
  geom_bar() +
  coord_flip() + 
  facet_grid(set ~ nbins)
```

```{r}
hamby224_test_explain %>%
  filter(set == "Set 1", nbins == "2") %>%
  mutate(rfscore = round(rfscore, 3)) %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  bind_rows(hamby224_test_explain %>%
                  filter(is.na(rfscore), set == "Set 1") %>%
                  select(case, bullet1, bullet2, land1, land2, rfscore)) %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
             text = paste('Bullets Compared: ', bullet1, "-", land1, 
                          "vs", bullet2, "-", land2,
                          '\nRandom Forest Score: ', 
                          ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet2 ~ bullet1, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(x = "", y = "", fill = "RF Score")
```

>>>>

```{r}
# Create a subsampled version of the training data
if (!file.exists("../data/hamby173and252_train_sub.csv")) {

  # Create the data (with a set seed)
  set.seed(1837592)
  hamby173and252_train_sub <- hamby173and252_train %>%
    filter(samesource == FALSE) %>%
    slice(sample(1:length(hamby173and252_train$barrel1), 4500, replace = FALSE)) %>%
    bind_rows(hamby173and252_train %>% filter(samesource == TRUE))

  # Save the subsampled data
  write.csv(hamby173and252_train_sub, "../data/hamby173and252_train_sub.csv", row.names = FALSE)

  # Load in the subsampled training data
  hamby173and252_train_sub <- read.csv("../data/hamby173and252_train_sub.csv")

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../data/hamby224_lime_inputs.rds")) {

  # Specify the options to use with lime
  hamby224_lime_inputs <- data.frame(data_name = c(rep("Full", 12), rep("Subsampled", 12)),
                                     bin_continuous = rep(c(rep(TRUE, 10), rep(FALSE, 2)), 2),
                                     quantile_bins = rep(c(rep(TRUE, 5), rep(FALSE, 5), rep(TRUE, 2)), 2),
                                     nbins = rep(c(rep(2:6, 2), rep(4, 2)), 2),
                                     use_density = rep(c(rep(TRUE, 10), TRUE, FALSE)), 2) %>%
    mutate(case = 1:length(data_name)) %>%
    select(case, data_name, bin_continuous, nbins, quantile_bins, use_density)

  # Apply lime to the full training data with the specified input options
  hamby224_lime_explain_full <- pmap(.l = as.list(hamby224_lime_inputs %>% 
                                                    filter(data_name == "Full") %>%
                                                    select(-case, -data_name)),
             .f = run_lime, # run_lime is one of my helper functions
             data_name = "Full",
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = TRUE)

  # Separate the lime and explain function results from the full data
  hamby224_lime_full <- map(hamby224_lime_explain_full, function(list) list$lime)
  hamby224_explain_full <- map_df(hamby224_lime_explain_full, function(list) list$explain)

  # Apply lime to the subsampled training data with the specified input options
  hamby224_lime_explain_sub <- pmap(.l = as.list(hamby224_lime_inputs %>% 
                                               filter(data_name == "Subsampled") %>%
                                               select(-case, -data_name)),
             .f = run_lime, # run_lime is one of my helper functions
             data_name = "Subsampled",
             train = hamby173and252_train_sub %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = TRUE)

  # Separate the lime and explain function results from the subsampled data
  hamby224_lime_sub <- map(hamby224_lime_explain_sub, function(list) list$lime)
  hamby224_explain_sub <- map_df(hamby224_lime_explain_sub, function(list) list$explain)

  # Join the lime results from the full and subsampled training data
  hamby224_lime <- c(hamby224_lime_full, hamby224_lime_sub)
  hamby224_explain <- bind_rows(hamby224_explain_full, hamby224_explain_sub)

  # Name the items in the lime list
  names(hamby224_lime) <- map_chr(1:dim(hamby224_lime_inputs)[1], function(case) 
      sprintf("case: data = %s, bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s",
              hamby224_lime_inputs$data_name[case],
              hamby224_lime_inputs$bin_continuous[case],
              hamby224_lime_inputs$quantile_bins[case],
              hamby224_lime_inputs$nbins[case],
              hamby224_lime_inputs$use_density[case]))

  # Save the lime objects
  saveRDS(hamby224_lime_inputs, "../data/hamby224_lime_inputs.rds")
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

@@

  # Load in the lime objects
  hamby224_lime_inputs <- readRDS("../data/hamby224_lime_inputs.rds")
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

@@

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = pmap_chr(list(feature = feature, 
                                  feature_value = feature_value,
                                  d_n = training_data, 
                                  b_c = bin_continuous,
                                  q_b = quantile_bins,
                                  n_b = nbins,
                                  u_d = use_density),
                            .f = bin_labeller, # bin_labeller is one of my helper functions
                            bin_data = hamby224_bin_boundaries,
                            case_info = hamby224_lime_inputs)) %>%
    mutate(feature = factor(feature),
           nbins = factor(nbins),
           feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    arrange(nbins)

@@

  # Reorder the variables of feature_desc and feature_bin for plotting purposes
  hamby224_test_explain <- hamby224_test_explain %>%
    mutate(feature_desc = reorder(feature_desc, strictly_less),
           feature_desc = reorder(feature_desc, feature_number),
           feature_desc = reorder(feature_desc, as.numeric(feature))) %>%
    select(training_data:use_density, case:feature_desc, feature_bin:strictly_less, 
           data, prediction)

@@

@@

  # Create a data frame with the interesting information relating to the differe
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  hamby224_lime_comparisons <- hamby224_test_explain %>%
    select(-data, -prediction) %>%
    group_by(training_data, case, bin_continuous, quantile_bins, nbins, use_density) %>%
    slice(1) %>%
    ungroup() %>%
    select(training_data, bin_continuous, quantile_bins, nbins, use_density, 
           set, case, rf_features, rfscore, model_prediction, model_r2) %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

@@

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(training_data, bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(training_data, set)
```

```{r}
hamby224_lime_results %>%
  group_by(training_data, set) %>%
  arrange(mse) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = mse, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid(training_data ~ set) + 
  theme_bw()
```

```{r}
hamby224_lime_results %>%
  group_by(training_data, set) %>%
  arrange(desc(ave_r2)) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = ave_r2, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid(training_data ~ set) + 
  theme_bw()
```

@@

```{r}
hamby224_test_explain %>%
  filter(training_data == "Full", set == "Set 1", nbins == "2") %>%
  ggplot(aes(x = feature_desc)) + 
  geom_bar() +
  coord_flip() + 
  facet_grid(set ~ .)
```

```{r}
hamby224_test_explain %>%
  filter(training_data == "Full", set == "Set 1", nbins == "2") %>%
  mutate(rfscore = round(rfscore, 3)) %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  bind_rows(hamby224_test_explain %>%
                  filter(is.na(rfscore), set == "Set 1") %>%
                  select(case, bullet1, bullet2, land1, land2, rfscore)) %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
             text = paste('Bullets Compared: ', bullet1, "-", land1, 
                          "vs", bullet2, "-", land2,
                          '\nRandom Forest Score: ', 
                          ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet2 ~ bullet1, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(x = "", y = "", fill = "RF Score")
```
